<!--
%\VignetteEngine{knitr}
%\VignetteIndexEntry{Running queries on a database}
-->

```{r, echo = FALSE, message = FALSE}
library(berdie)
```

# Loading a database connection

Database queries produce results in table format -- a certain number of rows by columns.
This is essentially a data.frame, so it makes sense to be able to turn queries into
data.frames given a database connection.

A database connection, like in Rails, is given by a database.yml file. You can pass in 
the file explicitly to `postgresql_connection`:

```{r, eval = FALSE}
connection <- postgresql_connection("some/dir/database.yml")
```

However, it is easier to let the package do its job and set either

- The `database.yml` option with the location of the configuration file.
  You can ensure this always runs on startup by placing it in your `~/.Rprofile` file
  with `options(database.yml = "some/dir/database.yml")`.

- Load the syberia package. If a syberia project is already loaded,
  the current value of `syberia_root()` will be used and the relative
  file `config/database.yml` will give the database configuration. Thus, if your
  syberia project is at `some/dir`, the default connection used by Berdie
  will be the one given by `some/dir/config/database.yml`.

# Running a query on a database

Once you have an open connection, running queries is straightforward:

```{r, eval = FALSE}
df <- run_query("select * from some_table limit 5", connection)
# A data.frame with (at most) five rows.
```

However, you do not have to lug the connection around.

```{r, eval = FALSE}
df <- run_query("select * from some_table limit 5")
```

The default is to use the helper function `last_connection()`, which will attempt
to look for: (1) the previous connection used for a query execution,
(2) load the connection from the `database.yml` global option, or (3) borrow
the database.yml file from a syberia project.

Whatever the case, you can always inspect the connection used by the previous query
by asking for `last_connection()`. If you place `library(syberia)` and `library(berdie)`
in your `~/.Rprofile` and you have set a default syberia project, you should be able
to execute `run_query` on starting R without any setup.

